---
title: "Alex alldata analyse as Notebook"
output: html_notebook
date: 12th May 2020 

---

Starting with script from from: https://osf.io/p3ygd/ called alldata_analyse, and working through in chunks to check I understand it.  

Script collates data from autistic and non-autistic adults, then runs confirmatory factor analyses to assess relationships between tests, and group differences in scores on the tests

```{r mylibs}
library(MVN)
library(psych)
library(rcompanion)
library(lavaan)
library(semPlot)
library(dplyr)
library(eeptools)
library(yarrr)
library(Hmisc) #needed for describe function
```
 COLLATING ALL DATA

Read in processed data for tasks (this data incorporates that from autistic and non-autistic people).  

This bit is modified by DB to read data direct from OSF

```{r collatedata}
impdata<-read.csv('https://osf.io/me5xq/download')
vocabdata<-read.csv('https://osf.io/52vuw/download')
grammardata<-read.csv('https://osf.io/t9xqe/download')
inferencingdata<-read.csv('https://osf.io/ru5nz/download')
backchanneldata<-read.csv('https://osf.io/tm5er/download')
TOMdata<-read.csv('https://osf.io/f2y8z/download')
awkwarddata<-read.csv('https://osf.io/yhr2c/download')

##Read in further variables
demographicsAut<-read.csv('https://osf.io/rhua2/download',stringsAsFactors = F)
demographicsNonAut<-read.csv('https://osf.io/7tx9b/download',stringsAsFactors = F)
IQdata<-read.csv('https://osf.io/euv62/download') 
ADOSdata<-read.csv('https://osf.io/xe7wm/download',stringsAsFactors=F)
CCSRdata<-read.csv('https://osf.io/2nz54/download',stringsAsFactors=F)
```


Make full data frame and add gender
```{r makedf}
alldata<-data.frame(impdata$ID,impdata$Group,
                    impdata$imp_total,
                    backchanneldata$total,awkwarddata$factor_scores,TOMdata$factor_scores,
                    grammardata$total,vocabdata$total,
                    inferencingdata$total,TOMdata$raw_CONTtotal,impdata$exp_acc)

colnames(alldata)<-c("ID","Group","implicature","fillers","awkwardconvo","TOManimations","grammar","vocab","inferencing","TOMcontrol","impcontrol")

##Get Gender variable for non-autistic adults
alldata$Gender<-NA
for(i in 1:120){
  myID<-impdata$ID[i]
  myID<-as.numeric(as.character(myID))
  
  if(myID %in% demographicsNonAut$ID){
    alldata$Gender[i]<-demographicsNonAut$gender[which(demographicsNonAut$ID==myID)]
  }
}

##Get Gender for autistic adults
for(i in 121:194){
  
  myID<-impdata$ID[i]
  myID<-as.numeric(as.character(myID))
  
  if(myID %in% demographicsAut$ID){
    alldata$Gender[i]<-demographicsAut$gender[which(demographicsAut$ID==myID)]
  }
}

```
 Also get IQ, ADOS and CC-SR data for autistic participants

```{r getiqetc}
alldata$IQ<-NA
alldata$ADOSscore<-NA
alldata$ADOSclass<-NA
alldata$CCSRpragRawscore<-NA

for(i in 121:length(alldata$ID)){
  
  myID<-alldata$ID[i]

  if(myID %in% ADOSdata$ID){
    alldata$ADOSscore[i]<-ADOSdata$TOTAL[which(ADOSdata$ID==myID)]
    alldata$ADOSclass[i]<-ADOSdata$Classification[which(ADOSdata$ID==myID)]
  }
  if(myID %in% CCSRdata$ID){
    alldata$CCSRpragRawscore[i]<-CCSRdata$Pragcorrected[which(CCSRdata$ID==myID)]
  }

  if(myID %in% IQdata$ID){
    alldata$IQ[i]<-IQdata$total[which(IQdata$ID==myID)]
  }
}

```


 Carry out exclusions

```{r exclusions}
#Check IDs all lined up. They are!
IDs<-data.frame(impdata$ID,backchanneldata$ID,awkwarddata$ID,TOMdata$ID,
                inferencingdata$ID,grammardata$ID,vocabdata$ID)

##Need to check inferencing totals for outliers
inf.desc.stats<-describe(alldata$inferencing,IQR=T,quant=c(.25,.75))
find.outliers<-inf.desc.stats$Q0.25-inf.desc.stats$IQR*2.2
outliers<-which(alldata$inferencing<find.outliers)
inf.outliers<-rep(0,194)
inf.outliers[outliers]<-1

##Recode univariate outliers as NA prior to analysis below
##(vocab, fillers and TOManimations have no outliers)
##Note that outliers can be inspected below under exclusions
alldata.outliers<-data.frame(impdata$outlier,grammardata$outlier,awkwarddata$outlier,inf.outliers)
alldata$implicature[which(alldata.outliers$impdata.outlier==1)]<-NA
alldata$impcontrol[which(alldata.outliers$impdata.outlier==1)]<-NA
alldata$grammar[which(alldata.outliers$grammardata.outlier==1)]<-NA
alldata$inferencing[which(alldata.outliers$inf.outliers==1)]<-NA
alldata$awkwardconvo[which(alldata.outliers$awkwarddata.outlier==1)]<-NA

#find autistic participants who did not have clinical diagnoses
exclude<-c(507125,507148,551069)
exclude.indices<-numeric()
for(i in 1:3){
exclude.indices<-c(exclude.indices,which(alldata$ID==exclude[i]))
}

rawdata<-alldata # keep record of all data, including excluded participants
alldata<-alldata[-exclude.indices,] # main data-set excluding 3 participants; autistic group includes anyone reporting a diagnosis
alldata1<-alldata[c(which(alldata$Group=="Non-autistic"),which(alldata$ADOSclass=="autism spectrum"),which(alldata$ADOSclass=="autism")),] # secondary data-set; autistic group only includes
#those meeting ADOS-2 criteria

str(alldata)
```

 MAKE TABLE OF DESCRIPTIVE STATS
 and also store info re exclusions

```{r makedesc}
all.desc.stats<-psych::describeBy(alldata[,c(3,11,4:6,10,7:9)],group=alldata$Group)
#print(all.desc.stats,signif=3)

##Calculate Cohen's d for all variables:
##In columns 1:3, place estimates and CIs for all autistic people compared to non-autistic
##In columns 4:6, place estimates and CIs for autistic people meeting ADOS-2 criteria compared to non-autistic people
cohen.d<-data.frame(rep(NA,8),rep(NA,8),rep(NA,8),rep(NA,8),rep(NA,8),rep(NA,8))
row<-1
for(i in 3:10){
cohen.d[row,1]<-effsize::cohen.d(d=alldata[,i],f=alldata$Group,na.rm=T)$estimate
cohen.d[row,2:3]<-effsize::cohen.d(d=alldata[,i],f=alldata$Group,na.rm=T)$conf.int
cohen.d[row,4]<-effsize::cohen.d(d=alldata1[,i],f=alldata1$Group,na.rm=T)$estimate
cohen.d[row,5:6]<-effsize::cohen.d(d=alldata1[,i],f=alldata1$Group,na.rm=T)$conf.int
row<-row+1
}
row.names(cohen.d)<-colnames(alldata)[3:10]
colnames(cohen.d)<-c("Cohen.d.full.sample","Lower95CI.1","Upper95CI.1","Cohen.d.AutMeetADOS","Lower95CI.1","Upper95CI.1")
print(cohen.d)

exclusions<-data.frame(c("implicature","Awkward Dialogues","Inferencing","Grammar"),rep(NA,4),rep(NA,4))
colnames(exclusions)<-c("Test","Outlying.value","Participant")
#Collect outlying values, and the 
exclusions[1,2]<-paste(impdata$imp_total[which(impdata$outlier==1)],collapse=" ")
exclusions[4,2]<-paste(grammardata$total[which(grammardata$outlier==1)],collapse=" ")
exclusions[3,2]<-paste(inferencingdata$total[which(inf.outliers==1)],collapse=" ")
exclusions[2,2]<-paste(round(awkwarddata$factor_scores[which(awkwarddata$outlier==1)],2),collapse=" ")
exclusions[1,3]<-paste(impdata$ID[which(impdata$outlier==1)],collapse=" ")
exclusions[4,3]<-paste(grammardata$ID[which(grammardata$outlier==1)],collapse=" ")
exclusions[3,3]<-paste(inferencingdata$ID[which(inf.outliers==1)],collapse=" ")
exclusions[2,3]<-paste(awkwarddata$ID[which(awkwarddata$outlier==1)],collapse=" ")
print(exclusions)
```



 CONFIRMATORY FACTOR ANALYSES

 PREPARE DATA


In selecting estimator for lavaan models, note that MLR estimator allows calculation of robust standard errors and test statistic for incomplete data. 

```{r inspect.corrs}
##Inspect non-transformed correlations
cor.matrix.notrans<-cor(alldata[,3:9],method="pearson",use="pairwise.complete.obs")
cor.matrix.notrans
```

##Check normality of variables.
```{r normcheck}
normplots<-mvn(na.omit(alldata[,3:9]), mvnTest = "royston", univariatePlot = "histogram")
normtest.nontrans<-mvn(na.omit(alldata[,3:9]), multivariateOutlierMethod = "adj",showOutliers=T)
```
Note that several cases are identified as multivariate outliers.
With transformations, none are identified, suggesting analysis should proceed with transformed data.

Let's take a look at the outliers in the non-transformed data
```{r view_outliers}
mvoutliers.nontrans.find<-as.numeric(normtest.nontrans$multivariateOutliers$Observation)
mvoutliers.nontrans<-alldata[mvoutliers.nontrans.find,]

##Do we want to transform?
transform<-1

if(transform==1){
  ##Add an arbitrary constant to TOM and awkward conversations data, as the
  ##transformation requires there be no negative numbers
  alldata$TOManimations<-alldata$TOManimations+3
  alldata$awkwardconvo<-alldata$awkwardconvo+3
  
  ##Use the Tukey transform -- this establishes the best power to which to raise each variable,
  ##so that it is more normally distributed. 
  for(i in 3:9){
    alldata[,i]<-transformTukey(alldata[,i],plotit=FALSE)
  }

  ##We need to change the scales of some variables so that they are more similar
  alldata[,c(3,4)]<-alldata[,c(3,4)]/100
  alldata[,c(7,9)]<-alldata[,c(7,9)]/100000

  #Now inspect transformed data
  cor.matrix.trans<-cor(alldata[,3:9],method="pearson",use="pairwise.complete.obs")
  normplots.trans<-mvn(na.omit(alldata[,3:9]), mvnTest = "royston", univariatePlot = "histogram")
  normtest.trans<-mvn(na.omit(alldata[,3:9]), multivariateOutlierMethod = "adj",showOutliers=T)
  mvoutliers.trans.find<-as.numeric(normtest.trans$multivariateOutliers$Observation)
}
```

# TEST HYPOTHESES


## HYPOTHESIS 1: Does a two-factor or one-factor model better account for the data across autistic and non-autistic individuals?

Specify models
```{r modelspec}
twofactor<-'
ONE=~implicature+fillers+awkwardconvo+TOManimations
TWO=~vocab+grammar
'

onefactor<-'
ONE=~implicature+fillers+vocab+grammar+awkwardconvo+TOManimations
'

##Now run two models

twofit<-cfa(model=twofactor,data=alldata,meanstructure=T,std.lv=T,missing="fiml",estimator="MLR")
two.fit.stats<-as.data.frame(fitMeasures(twofit,c("cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper")))
#There is a warning hear about negative variances (i.e. there is a Heywood case)
#This is an indication that there is model mis-specification
#Inspecting correlations above, looks like this is due to grammar clustering
#with factor ONE

onefit<-cfa(model=onefactor,data=alldata,meanstructure=T,std.lv=T,missing="fiml",estimator="MLR") 
one.fit.stats<-as.data.frame(fitMeasures(onefit,c("cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper")))
#One factor model is weak
```
Define new model without vocab, as this variable seems distinct from other language tests

```{r revised_model}
revised<-'
COMM=~implicature+fillers+grammar+awkwardconvo+TOManimations+inferencing
'
fit.revised<-cfa(model=revised,data=alldata,meanstructure=T,std.lv=T,missing="fiml",estimator="MLR") 
fit.revised.stats<-as.data.frame(fitMeasures(fit.revised,c("cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper")))
colnames(fit.revised.stats)[1]<-'index'
##Good fit stats!
fit.revised
fit.revised.stats
##Store factor scores for this model
alldata<-data.frame(alldata,lavPredict(fit.revised))
```
## HYPOTHESIS 2: Are there significant group differences at the level of the factors?

Now add Group as a covariate to the model, and compare models where path between Group (i.e. autistic or non-autistic) and the factors is allowed to vary and where it isn't
```{r testH2}
with.Group.effect<-'
ONE=~implicature+fillers+awkwardconvo+TOManimations+grammar+inferencing
ONE~Group
'
no.Group.effect<-'
ONE=~implicature+fillers+awkwardconvo+TOManimations+grammar+inferencing
ONE~0*Group
'

fit.with.Group<-cfa(model=with.Group.effect,data=alldata,meanstructure=T,std.lv=T,missing="fiml",estimator="MLR")
Group.fit.stats<-as.data.frame(fitMeasures(fit.with.Group,c("cfi.robust","rmsea.robust","rmsea.ci.lower.robust","rmsea.ci.upper.robust")))
colnames(Group.fit.stats)[1]<-'index'
Group.fit.stats
fit.no.Group<-cfa(model=no.Group.effect,data=alldata,meanstructure=T,std.lv=T,missing="fiml",estimator="MLR")
no.Group.fit.stats<-as.data.frame(fitMeasures(fit.no.Group,c("cfi.robust","rmsea.robust","rmsea.ci.lower.robust","rmsea.ci.upper.robust")))
colnames(no.Group.fit.stats)[1]<-'index'
no.Group.fit.stats
chidiff.group.effect<-anova(fit.with.Group,fit.no.Group)
chidiff.group.effect
```

Clearly there are group differences, as model with the group effect fixed to zero is very weak. However, model allowing a group difference could have
 better fit .... so investigate whether the residuals for particular test(s) are contributing to problems with model fit.

```{r checkpathgroupmodel}
resid<-residuals(fit.with.Group,type="cor")$cov
##Looks like we need paths between Group and imp too

with.Group.effectV2<-'
ONE=~implicature+fillers+awkwardconvo+TOManimations+grammar+inferencing
ONE+implicature~Group
'
revised.Group.fit<-cfa(model=with.Group.effectV2,data=alldata,meanstructure=T,std.lv=T,estimator="MLR",missing="fiml")
revised.Group.fit.stats<-as.data.frame(fitMeasures(revised.Group.fit,c("cfi.robust","rmsea.robust","rmsea.ci.lower.robust","rmsea.ci.upper.robust")))
revised.Group.result<-standardizedSolution(revised.Group.fit)
revised.Group.result
```

Test for gender differences  

There is group x gender interaction
```{r genderdiffs}
alldata$Gender[alldata$Gender=="Female "]<-"Female"
alldata$Gender[alldata$Gender==""]<-NA
Gender.diffs<-summary(aov(alldata$COMM~alldata$Gender*alldata$Group,data=alldata))
Gender.diffs
```

HYPOTHESIS 3: Are factor scores linked to self-reported and observed difficulties in social interaction in autistic individuals?

```{r testH3}
variability.communication<-'
COMM=~implicature+fillers+awkwardconvo+TOManimations+grammar+inferencing
ADOSscore+CCSRpragRawscore~COMM
'

fit.variability.communication<-cfa(model=variability.communication,data=alldata[alldata$Group=="Autistic",],meanstructure=T,std.lv=T,estimator="MLR",missing="fiml")
variability.communication.fit.stats<-as.data.frame(fitMeasures(fit.variability.communication,c("cfi.robust","rmsea.robust","rmsea.ci.lower.robust","rmsea.ci.upper.robust")))
variability.communication.results<-standardizedSolution(fit.variability.communication)
#YES! Moderate pathways of about .4
variability.communication.results
```


Can also check extent to which links between factor scores and other measures
remain when correcting for general intelligence -- these regressions allow extraction of IQ-corrected residuals

```{r IQresids}
alldata$ADOSresid<-resid(lm(alldata$ADOSscore ~ alldata$IQ, na.action = na.exclude))
alldata$CCSRresid<-resid(lm(alldata$CCSRpragRawscore ~ alldata$IQ, na.action = na.exclude))
```

Make plots

```{r plotGroupFA}
##Make path diagram of factor analysis comparing groups
x<-c(-2,-1.5,-1,-0.5,0,0.5,-2,-0.75)
y<-c(0,0,0,0,0,0,0.6,1)
lay<-matrix(c(x,y),ncol=2)
semPlot::semPaths(revised.Group.fit,"std",intercepts=F,layout=lay,nodeLabels = c("imp","flr","awk","TOM","grm","inf","Diag","COMM"))
```
Make path diagram of factor analysis exploring variability in autistic group

```{r autvarplot}
x<-c(-2,-1.5,-1,-0.5,0,0.5,-1.5,0,-0.75)
y<-c(0,0,0,0,0,0,-2,-2,-1)
lay<-matrix(c(x,y),ncol=2)
semPlot::semPaths(fit.variability.communication,"std",intercepts=F,layout=lay,nodeLabels = c("imp","flr","awk","TOM","grm","inf","ADOS","CCSR","COMM"))

```
Now make pirate plot showing factor scores in two groups by Gender

Standardise factor scores to have SD of 1 across the sample to allow easier interpretation of group differences in pirate plot

```{r dopirateplot}
alldata$COMM<-alldata$COMM/sd(alldata$COMM)
pirateplot(COMM ~ Gender*Group, data=alldata,
           ylab="Receptive communication factor score",
           pal="gray",point.o=1,inf.method="ci")

#I removed yaxt.y=seq(-2,3,1), as this caused chunk to stall - possible a different version of yarrr?
```

## Final check
Check whether the factor analysis measured communication skills in a similar way across groups through measurement invariance testing.  

We need to set up models in which we progressively fix parameters. Note that scores on the implicature test showed a group difference, even accounting for factor level differences, so we need to leave the intercept on that test to vary.

# Model with no constraints
```{r m.noconstraints}
fit.group.noconstraints<-cfa(model=revised,data=alldata,group="Group",
                                std.lv=T,missing="fiml",estimator="MLR")
fit.stats.no.constraints<-t(data.frame(fitMeasures(fit.group.noconstraints,c("cfi.robust","rmsea.robust"))))
fit.stats.no.constraints
```
# Model constraining loadings to be the same across groups
```{r sameloadings}
fit.group.loadings<-cfa(model=revised,data=alldata,group="Group",group.equal=c("loadings"),
                        std.lv=T,missing="fiml",estimator="MLR")
fit.stats.loadings<-t(data.frame(fitMeasures(fit.group.loadings,c("cfi.robust","rmsea.robust"))))
fit.stats.loadings

```
# Model constraining intercepts to be the same across groups
```{r sameintercepts}
fit.group.intercepts<-cfa(model=revised,data=alldata,group="Group",group.equal=c("loadings","intercepts"),group.partial=c("implicature~1"),
                       std.lv=T,missing="fiml",estimator="MLR")
fit.stats.intercepts<-t(data.frame(fitMeasures(fit.group.intercepts,c("cfi.robust","rmsea.robust"))))
fit.stats.intercepts
```

# Model constraining residuals to be the same across groups
```{r sameresids}
fit.group.residuals<-cfa(model=revised,data=alldata,group="Group",group.equal=c("loadings","intercepts","residuals"),group.partial=c("implicature~1"),
                         std.lv=T,missing="fiml",estimator="MLR")
fit.stats.residuals<-t(data.frame(fitMeasures(fit.group.residuals,c("cfi.robust","rmsea.robust"))))
fit.stats.residuals
```

# Now test for difference between models with and without constraints
```{r modelcompare}
anova.test.measInv.metric<-as.data.frame(anova(fit.group.noconstraints,fit.group.loadings))
anova.test.measInv.metric
anova.test.measInv.scalar<-as.data.frame(anova(fit.group.loadings,fit.group.intercepts))
anova.test.measInv.scalar
anova.test.measInv.strict<-as.data.frame(anova(fit.group.intercepts,fit.group.residuals))
anova.test.measInv.strict
summary(fit.group.residuals,fit.measures=T)

```

If we look at the most constrained model, we see that fit stats are high, indicating that the model fits equally well in both groups. The only parameters left to vary are means and variances of the factors. Autistic group has lower mean and bigger variance
